### 2023-12-28 els2024-convex-covering/paper.tex
Scaffolding

\documentclass[format=sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{float}
\usepackage[labelfont=bf,textfont=md]{caption}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{hyperref}
\usepackage[parfill]{parskip}
\usepackage[all]{hypcap}
\usemintedstyle[common-lisp]{default}
\newmintinline[code]{text}{}
\bibliographystyle{plainnat}

\hypersetup{
  colorlinks,
  linkcolor={red!50!black},
  citecolor={blue!50!black},
  urlcolor={blue!80!black}
}

\newlist{step}{enumerate}{10}
\setlist[step]{label*=\arabic*.,leftmargin=2em}

\acmConference[ELS’24]{the 17th European Lisp Symposium}
{May 6--7 2024}{Vienna, Austria}
\acmDOI{}
\setcopyright{rightsretained}
\copyrightyear{2024}

\begin{document}

\title{Paper: Convex Covering}

\author{Jan ``scymtym'' Moringen}
\email{jmoringe@techfak.uni-bielefeld.de}
\author{Yukari ``Shinmera'' Hafner}
\email{shinmera@tymoon.eu}
\affiliation{%
  \institution{Shirakumo.org}
  \city{Zürich}
  \country{Switzerland}
}

\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10010147.10010371.10010396.10010398</concept_id>
  <concept_desc>Computing methodologies~Mesh geometry models</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10010147.10010371</concept_id>
  <concept_desc>Computing methodologies~Computer graphics</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10010405.10010469.10010474</concept_id>
  <concept_desc>Applied computing~Media arts</concept_desc>
  <concept_significance>100</concept_significance>
  </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Mesh geometry models}
\ccsdesc[500]{Computing methodologies~Computer graphics}
\ccsdesc[100]{Applied computing~Media arts}

\begin{abstract}
  A
\end{abstract}

\keywords{Common Lisp, Convex Decomposition, Games, Video Games, Computer Graphics, Experience Report}

\maketitle

\def\abovecaptionskip{1pt}
\def\listingautorefname{Listing}
\def\figureautorefname{Figure}

\section{Introduction}\label{introduction}

\section{Related Work}\label{relatedwork}

Liu et al.\cite{liu2008convex}

\section{Algorithm}\label{algorithm}

\section{Extensions}\label{extensions}

\section{Conclusion}\label{conclusion}

\section{Further Work}\label{further-work}

\section{Acknowledgements}\label{acknowledgements}

\bibliography{paper}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-command-extra-options: "-shell-escape"
%%% TeX-master: t
%%% TeX-engine: luatex
%%% End: