### 2025-01-31 els2025-sbcl-nx/paper.tex
Edit build section and add section on relocation.

\documentclass[format=sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{enumitem}
\usepackage{float}
\usepackage[labelfont=bf,textfont=md]{caption}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{hyperref}
\usepackage[parfill]{parskip}
\usepackage[all]{hypcap}
\usemintedstyle[common-lisp]{default}
\newmintinline[code]{text}{}
\bibliographystyle{plainnat}

\hypersetup{
  colorlinks,
  linkcolor={red!50!black},
  citecolor={blue!50!black},
  urlcolor={blue!80!black}
}

\newlist{step}{enumerate}{10}
\setlist[step]{label*=\arabic*.,leftmargin=2em}

\acmConference[ELS’25]{the 18th European Lisp Symposium}
{May 19--20 2025}{Zürich, Switzerland}
\acmDOI{}
\setcopyright{rightsretained}
\copyrightyear{2025}

\begin{document}

\title{Porting the Steel Bank Common Lisp Compiler and Runtime to the Nintendo Switch NX Platform}

\author{Charles Zhang}
\email{charleszhang99@yahoo.com}
\author{Yukari ``Shinmera'' Hafner}
\email{shinmera@tymoon.eu}
\affiliation{%
  \institution{Shirakumo.org}
  \city{Zürich}
  \country{Switzerland}
}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10011007.10011006.10011041.10011048</concept_id>
       <concept_desc>Software and its engineering~Runtime environments</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011006.10011041.10011045</concept_id>
       <concept_desc>Software and its engineering~Dynamic compilers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10010940.10010941.10010949.10010950.10010954</concept_id>
       <concept_desc>Software and its engineering~Garbage collection</concept_desc>
       <concept_significance>300</concept_significance>
       </concept>
   <concept>
       <concept_id>10011007.10011074</concept_id>
       <concept_desc>Software and its engineering~Software creation and management</concept_desc>
       <concept_significance>100</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Runtime environments}
\ccsdesc[500]{Software and its engineering~Dynamic compilers}
\ccsdesc[300]{Software and its engineering~Garbage collection}
\ccsdesc[100]{Software and its engineering~Software creation and management}

\begin{abstract}
  The Nintendo Switch (NX) is a 64-bit ARM-based platform for video games with a proprietary micro-kernel operating system. Notably this system does not give programs the ability to mark pages as executable at load time or give access to thread signal handlers, both of which present a significant hurdle to SBCL's intended bootstrap process and runtime operation. We present our efforts to adapt the SBCL runtime and compiler to deploy applications onto the NX platform.
\end{abstract}

\keywords{Common Lisp, SBCL, porting, ARM, aarch64, NX, Experience Report}

\maketitle

\def\abovecaptionskip{1pt}
\def\listingautorefname{Listing}
\def\figureautorefname{Figure}

\section{Introduction}\label{introduction}
The Nintendo Switch (codename NX) is a handheld video game console based on the ARM 4 Cortex-A57 64-bit architecture\cite{morgan2020}. It features a proprietary micro-kernel operating system called the ``Nintendo Switch system software''\cite{roussel2019methodically} (NX OS), and normally only runs software that has been licensed, approved, and digitally signed and encrypted by Nintendo.

Developing licensed software for the NX has to be done via Nintendo's own proprietary SDK, which they only distribute under a non-disclosure agreement (NDA). An open-source third-party alternative to the SDK is available\cite{switchbrew}, but this cannot be used for licensed software.

The OS-provided runtime environment is only directly suitable for C and C++ software. Other game engines such as Unity, Unreal, and Godot do provide exporting functionality to the NX, meaning that ports for the runtime environments they rely on such as C\#, Lua, and GDScript have been developed, but remain closed-source.

Many high-performance native-code Common Lisp implementations operate by compiling and loading user code at runtime into the same process (the state of which is traditionally termed an `image' in the Lisp world). These implementations typically provide a way to dump an image in a way where the runtime of a new process can load it. This contrasts with other languages such as C where code is compiled and linked `offline' into an executable whose entire code and runtime gets loaded into its own process by the operating system. We cannot compile and load new code at runtime after an executable is loaded by the operating system on the NX due to platform security restrictions. Because large portions of the system such as the compiler of such image-based Common Lisp implementations are also written in Common Lisp, this means that the implementation must be bootstrapped and all user code compiled ahead of time off of the NX, before the compiled Lisp code in the image can then somehow be linked into a runtime built for the NX by the operating system, rather than relying on the runtime to load the image. We discuss the intricacies of this for an image-based implementation such as SBCL in \autoref{build} and \autoref{relocation}.

A port of a Common Lisp implementation such as ECL\cite{attardi1994embeddable}, where Lisp code is compiled to C and easily linked into a C runtime, is conceivable and presents far fewer challenges for building applications for the NX than with an image-based implementation. However, due to the high performance requirements of video games and the relatively low-power platform of the NX, we decided to try and port SBCL, since SBCL produces much faster code than ECL. In addition, some of the obstacles presented by the NX apply in general to how most Common Lisp implementations function.

For example, the NX OS does not provide user signal handlers. This is a problem for the Garbage Collector (GC), as SBCL relies on inter-thread signalling to park threads during garbage collection. While SBCL does provide a safepoints mechanism for GC that is used on Windows, this mechanism is not well tested on other platforms, and as-is still did not meet the requirements of this locked-down platform. We discuss the GC in detail in \autoref{gc}.

While we can now compile and deploy complex Common Lisp applications to the NX, some parts of the runtime remain unsupported, and we discuss our future efforts in this regard in \autoref{further-work}.

\section{Related Work}\label{relatedwork}
Rhodes\cite{rhodes2008sbcl} outlines the methodology behind the SBCL bootstrapping process.

A Common Lisp bootstrapping process as Durand et al.\cite{durand2019bootstrapping} outline where the whole target image is created on the host would work to avoid requiring compiling and executing code at runtime on the NX, however, there is not a complete implementation of this technique yet. User libraries which require calling foreign functions at compile time dependent on the target architecture also cannot be handled with this technique.

Citing information on the NX' operating system is difficult as it is a closed-source platform with all usual information placed under NDA. All publicly available information is from security research such as by Roussel-Tarbouriech et al.\cite{roussel2019methodically} and reverse-engineering\cite{switchbrew}.

Particularly, we are unaware of any publication about the porting of other runtime environments to the NX, such as C\#, JavaScript, Lua, etc.

Patton\cite{patton2023parallel} and Schwartz\cite{schwartz2018dynamic} describe some details about SBCL's garbage collector which we modify for our port.

\section{Build System}\label{build}
The usual SBCL build process proceeds in several distinct phases, some of which need to be run on a ``host system'' and others on the ``target system''. This does allow for limited cross-compilation, wherein the ``host system'' steps can be run on an operating system and platform that is not that of the target we are trying to compile for. However, the ``target system'' steps are supposed to run on the target architecture, and involve compiling, loading, and executing new system code in an existing image. As mentioned in the Introduction, we cannot load and execute code in the same process that compiled the code on the NX, as the NX OS does not allow us to map new executable pages.

The basic idea of our solution to this issue is to replace the NX as \textit{initial} target with an ARM64-based Linux. Once everything, including user code, has been compiled on this Linux target, we extract all Lisp code and data out using a process called \textit{shrinkwrapping} and link it together with the SBCL C runtime as compiled for the NX. This process ultimately results in a fully static ELF executable that does not perform any dynamic memory mapping or compilation.

Our build still functions largely the same as the one outlined by Rhodes\cite{rhodes2008sbcl}, though being run roughly twice with special configurations to accomplish the hybrid build.

\begin{enumerate}
\item \code{build-config} (NX) \\
  This step gathers whatever build configuration options for the target and spits them out into a readable format for the rest of the build process. We run this on some host system (which may not be our ARM64 Linux intermediary), using a special flag to indicate that we're building for the NX. We also enable the fasteval contrib, which we need to step in for any place where we would usually invoke the compiler at runtime.
\item \code{make-host-1} (NX) \\
  Next we build the cross-compiler with the host Lisp compiler, and at the same time emit C header files describing Lisp object layouts in memory as C structs for the next step.
\item \code{make-target-1} (NX) \\
  Now we use the C compiler the Nintendo SDK provides for us, which can cross-compile the SBCL C runtime for the NX. We had to make adjustments to the C runtime bits, as the NX OS is not POSIX compliant and lacks a few features the SBCL C runtime usually takes advantage of. The SBCL C runtime contains the GC and OS glue bits that the Lisp code needs.
\item \code{build-config} (Linux$\sim$NX) \\
  We now create an almost-normal ARM64 Linux system with the same feature set as for the NX. This involves the usual steps as before, though with a special flag to inform some parts of the Lisp process that we're going to ultimately target the NX.
\item \code{make-host-1} (Linux$\sim$NX)
\item \code{make-target-1} (Linux$\sim$NX)
\item \code{make-host-2} (Linux$\sim$NX)
  With the target runtime built, we build the target Lisp system (compiler and the standard library) using the Lisp cross-compiler built by the Lisp host compiler in \code{make-host-1} (Linux). This step produces a "cold core" that the runtime can jump into, and can be done purely on the host machine. This cold core is not complete, and needs to be executed on the target machine with the target runtime to finish bootstrapping, notably to initialise the object system, which requires runtime compilation.
\item \code{make-target-2} (Linux$\sim$NX) \\
  The cold core produced in the last step is loaded into the target runtime, and finishes the bootstrapping procedure to compile and load the rest of the Lisp system. After the Lisp system is loaded into memory, the memory is dumped out into a "warm core", which can be loaded back into memory in a new process with the target runtime. From this point on, new code can be loaded and images can be dumped at will.
\item \code{user} \\
  For user code we now perform some tricks to make it think it's running on the NX, rather than on Linux. In particular we modify \code{*features*} to include \code{:nx} and not \code{:linux}, \code{:unix}, or \code{:posix}. Once that is set up and ASDF has been neutered, we can compile our program "as usual" and at the end dump out a new core.
\item \code{shrinkwrap} \\
  Once all our code has been loaded and a final core has been produced, we ``shrinkwrap'' the image to produce assembly files. The details of this are outlined in \autoref{relocation}. These assembly files can then be linked together with the SBCL C runtime compiled for the NX in step (3) with the Nintendo SDK toolchain, producing a final ELF executable.
\item \code{package} \\
  The final step is to run all the other SDK toolchain bits to produce a signed application package, which can be deployed to a Nintendo Switch development kit to be run.
\end{enumerate}

A notable extra wrinkle in this procedure is that the SDK is only available for Windows. To accommodate this, the custom build system we developed can be run either directly from Windows using another ARM machine remotely to run the Linux bits, or it can be run from a Linux system with the ARM bits being run either locally or remotely and the Windows bits being run either remotely or through Wine\cite{amstadt1994wine}.

\section{Relocation}\label{relocation}
Lisp objects contain absolute pointers, so that the dumped image, which contains the memory state of the heap and other spaces, is filled with absolute pointers. Traditionally, image-based Lisp implementations such as SBCL map its address spaces to fixed addresses for two reasons: reloading an image into a process is easier this way, since objects can be mapped back in the new image directly to where they were in the process that dumped the image. In addition, some architecture backends optimize code generation by hardwiring the address of certain Lisp objects directly into the machine code.

Because the NX operating system enforces address space layout randomization, the Lisp runtime cannot rely on being able to map its address spaces to fixed address spaces. Hence, the runtime must relocate all objects present in the dumped image to the address spaces placed randomly by the operating space. However, code objects present a special problem on the NX. Code objects in SBCL contain both executable machine code and absolute pointers to Lisp objects residing in the heap. However, executable pages are not writable and must be mapped in by the operating system from the ELF core. Therefore, the code and data need to be separated, at first so that the runtime can relocate the absolute pointers to the Lisp objects, and later so that the moving garbage collector is able to fix up those references. This requires the shrinkwrapping procedure to separate the data from the code and to rewrite the code in the image offline so that the references to the pointers are into a dislocated read/write space instead of in the code object header, which is on an executable page and hence not writable on the NX. The code objects are put into code sections in the executable so that the operating system can load the code and then the runtime can later fix up references to these code objects in non-code objects residing in read/write space.

\section{Garbage Collection}\label{gc}
SBCL' default garbage collector, ``gencgc'', is a stop-the-world collector. In order to start collecting, it needs to ``park'' all threads in a safe location where they won't access any memory that the collector might move or change during collection.

On Unix systems signals are used to park the threads. By sending a signal to every other thread, each thread will enter a signal handler, which can then park the thread. This is advantageous, since it leverages an operating system mechanism to interrupt thread execution and frees the thread from checking whether it should park.

On Windows no equivalent mechanism is available, so a technique called ``safepoints'' is employed instead. The compiler injects code at strategic locations, such as at function call boundaries and loop returns, which checks whether the thread should park explicitly. This strategy is notably slower than using signals, since the safepoints are not free, and is finicky, as the compiler could inject too many safepoints (slowing execution of user code significantly) or too few safepoints (making threads take too long to park).

In order to make the safepoint test as small and cheap as possible SBCL does employ another trick on Windows: each safepoint simply writes to a memory region. When GC is triggered, this region is marked as protected, causing any thread that tries to write to it to trap, which SBCL can intercept to park the thread instead.

On the NX we cannot install trap handlers, so the current safepoint strategy fails.

\colorbox{red}{
  \textbf{TODO CHARLES}: Add some details here.
}

\section{Conclusion}\label{conclusion}
We have demonstrated that it is possible to port SBCL even to a very restrictive platform that is hostile towards dynamic runtimes such as used for Lisp.

\begin{figure}[h]
  \centering
  \includegraphics[width=\linewidth]{photo.png}
  \caption{A photo of a demo running in the Trial Common Lisp game engine on the Nintendo Switch hardware.}
\end{figure}

Using a substitute build host that is similar enough to the target platform we can compile all code ahead of time, then substitute the base runtime and rewrite the resulting code in order to create an executable suitable for the target platform.

We have also managed to update the garbage collector to work on operating systems with more restrictive feature sets than are found on a typical desktop operating system.

\section{Further Work}\label{further-work}
Currently we rely on the fasteval system to circumvent runtime compilation restrictions. This is especially vital for CLOS, since the discriminating function is only compiled on first call of a generic function. However, since we know that we don't dynamically add or remove methods, we should be able to pre-compile all of these functions as well. We'd like to add such a CLOS freeze step to the build pipeline.

Callbacks from C to Lisp are currently left unimplemented. Callbacks in SBCL are implemented in part as hand-written assembly stubs, and are very hard to maintain. The current stubs for ARM64 do not work correctly in the presence of the shrinkwrapping steps that we make use of.

Finally there are a number of improvements we've made that we would like to upstream to lessen the maintenance burden. We've already contributed a bunch of the changes back upstream, but a lot of it is also tied to the proprietary SDK and cannot be open-sourced due to the NDA, and some other changes are too contentious for the rest of the SBCL team to want to upstream.

\section{Acknowledgements}\label{acknowledgements}
We would like to thank Douglas Katzman for his help and advice for various parts of the porting effort, as well as the rest of the SBCL maintenance team for their continuous improvements to the SBCL platform.

\bibliography{paper}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-command-extra-options: "-shell-escape"
%%% TeX-master: t
%%% TeX-engine: luatex
%%% End: