### 2025-04-25 els2025-sbcl-nx/presentation.tex
Start working on the presentation

\documentclass[14pt,t,aspectratio=169]{beamer}
\usepackage{fontspec}
\usepackage{color}
\usepackage{minted}
\usepackage{array}
\usepackage{emoji}

%% These fonts are non-free.
%% Comment out the lines if you don't have them.
\setmainfont{Equity Text A}
\setsansfont{Concourse T3}
\setmonofont{Triplicate T4}

\definecolor{bgcolor}{RGB}{255,240,240}
\definecolor{red}{RGB}{224,35,35}
\setbeamercolor{background canvas}{bg=bgcolor}
\setbeamercolor{normal text}{fg=black}
\setbeamercolor{titlelike}{fg=black}
\setbeamercolor{itemize item}{fg=red}
\setbeamercolor{enumerate item}{fg=red}
\setbeamertemplate{itemize items}[circle]
\setbeamertemplate{navigation symbols}{}
\usemintedstyle{monokai}
\newminted[lispcode]{common-lisp}{fontsize=\footnotesize}
\newminted[smalllispcode]{common-lisp}{fontsize=\scriptsize}
\def\code#1{{\color{codecolor}\texttt{#1}}}
\renewcommand{\theFancyVerbLine}{\color{darkgray}\large \oldstylenums{\arabic{FancyVerbLine}}}

\usebackgroundtemplate{\includegraphics[width=\paperwidth,height=\paperheight]{background}}
\begin{document}
{\usebackgroundtemplate{\includegraphics[width=\paperwidth,height=\paperheight]{firstpage}}
  \begin{frame}
    \color{white}
    \vspace{3cm}
    {\hspace{1.4cm} \LARGE Porting SBCL to the Nintendo Switch} \\
    \vspace{1cm}
    {\hspace{1.6cm} Yukari Hafner, Charles Zhang} \\
    \vspace{0.2cm}
    {\hspace{2.1cm}\texttt https://shirakumo.org}
  \end{frame}}

\begin{frame}
  \frametitle{The Device}
  \includegraphics[height=5cm]{switch}
  \begin{itemize}
  \item CPU: ARM 4 Cortex-A57 64-bit
  \item OS: ``Horizon OS'', proprietary micro-kernel
  \item SDK: C++, proprietary version of Clang
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Immediate Challenges}
  \begin{itemize}
  \item Everything is proprietary and under NDA \\
    \Rightarrow Scarce public information
  \item The OS is not BSD or even fully POSIX \\
    \Rightarrow Need new OS abstractions
  \item There are no inter-thread signals \\
    \Rightarrow Can't use usual GC tricks
  \item We are not allowed to create executable pages \\
    \Rightarrow No compilation at runtime
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Basic Ideas}
  \begin{itemize}
    \color{lightgray}
  \item Everything is proprietary and under NDA \\
    \textcolor{red}{\Rightarrow} \textcolor{black}{Only publicise our own interfaces}
  \item The OS is not BSD or even fully POSIX \\
    \textcolor{red}{\Rightarrow} \textcolor{black}{Write C(++) shim libraries for access}
  \item There are no inter-thread signals \\
    \textcolor{red}{\Rightarrow} \textcolor{black}{Use safepoints}
  \item We are not allowed to create executable pages \\
    \textcolor{red}{\Rightarrow} \textcolor{black}{Compile everything on linux and shrinkwrap}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Build Procedure}
  \begin{itemize}
  \item 
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Relocation}
  \begin{itemize}
  \item 
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Garbage Collection}
  \begin{itemize}
  \item 
  \end{itemize}
\end{frame}

\begin{frame}[c]{ }
  \centering
  {\Huge Live Demo}
\end{frame}

\begin{frame}
  \frametitle{Further Work}
  \begin{itemize}
  \item Optimising CLOS dispatch ahead of time \\
    \Rightarrow Christophe? \emoji{pleading-face}
  \item Optimising Trial and Kandria \\
    \Rightarrow Lots of profiling work that can be done on PC
  \item Porting to the Nintendo Switch 2 \\
    \Rightarrow As soon as plebians like us get access from almighty \\\quad Nintendo
  \end{itemize}
\end{frame}

{\usebackgroundtemplate{\includegraphics[width=\paperwidth,height=\paperheight]{firstpage}}
  \begin{frame}[c]{ }
    \centering\color{white}
    \vspace{2cm}
    {\LARGE Thank you!} \\
    \vspace{0.2cm}
    Consider supporting our work: \\
    \includegraphics[width=3cm]{patreon} \\
    {\texttt https://patreon.com/shinmera}
  \end{frame}}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% TeX-engine: luatex
%%% TeX-command-extra-options: "-shell-escape"
%%% End: